Missing the Point\(\colon\) Non-Convergence in Iterative Imputation Procedures

### 2020-04-22 3.Thesis/1.SimulationStudy/Writeup/manuscript.Rmd
Compute rhat and ac for pathological non-convergence case. Adjust functions by adding wrapper for general convergence function

**Move this after notation??** Figure 1 provides an overview of the steps involved with MI---from incomplete data, to $M$ completed datasets, to $M$ estimated quantities of interest ($\hat{Q}$s), to a single pooled estimate $\bar{Q}$. Missing data in $y$ is 'imputed' (i.e., filled in) $M$ times. The imputed data ($y_{imp}$) is combined with the observed data ($y_{obs}$) to create $M$ completed datasets. On each completed dataset, the analysis of scientific interest (the 'substantive model', or 'complete data model') is performed. The quantity of scientific interest (e.g., a regression coefficient) is denoted with $Q$. Since $Q$ is estimated on each completed dataset, $M$ separate $\hat{Q}$-values are obtained. These $M$ values are combined into a single pooled estimate $\bar{Q}$. **ALSO: make clear that iteration happens within the first step.**

```{r, out.width="\\linewidth", include=TRUE, fig.align="center", fig.cap="Scheme of the main steps in multiple imputation."}
knitr::include_graphics("./images/diagram.pdf") #adapted from [@buur18] par. 1.4.1.
```

@@

Main question: How can non-convergence be diagnosed? Sub-questions: Are common MCMC non-convergence diagnostics appropriate for MICE? And if so, which threshold should be used to diagnose non-convergence? How many iterations are sufficient/needed to be able to diagnose non-convergence? Are the default number of iterations sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)? **(These are part of the methods, not RQ per se: What are the effects of continued iterations on estimates, predictions and inferences? And finally, do the answers differ with varying missingness proportions? That is, we vary the nr of iterations and the missingness proportion because we assume that the algorithm has not reached convergence at $t=1$, and performs worse with moderate to high missingness (not so much with little or a LOT of missingness).)**

@@

The two challenges for convergence of iterative algorithms are mixing and stationarity [@gelm13], see Figure XYZ. Without mixing, imputation chains do not intermingle nicely. Without stationarity, there is trending within imputation chains. While the current practice is to inspect traceplots for signs of non-mixing and non-stationarity, diagnostic evaluation of these two challenges would be preferred. Non-stationarity within chains may be diagnosed with e.g., autocorrelation [$AC$; @scha97; @gelm13], numeric standard error ['MC error'; @gewe92], or Raftery and Lewis's [-@raft91] procedure to determine the effect of trending on the precision of estimates. A widely used diagnostic to monitor mixing between chains is the potential scale reduction factor $\widehat{R}$ ['Gelman-Rubin statistic'; @gelm92]. With a recently proposed adaptation, $\widehat{R}$ might also serve to diagnose non-stationarity, but this has not yet been thoroughly investigated [@veht19]. Therefore, mixing and stationarity will be evaluated separately in this study. As recommended [e.g., @cowl96, p. 898], $AC$ and $\widehat{R}$ will be used.

```{r, out.width='.49\\linewidth', fig.width=3.5, fig.height=3,fig.show='hold',fig.cap="Pathological non-convergence versus typical convergence."}
load("./images/nonc.Rdata")
load("./images/conv.Rdata")
par(mar = c(2, 2, 0.1, 0.1))
plot(nonc)
plot(conv)

@@

Research question when and how to diagnose non convergence?

QUESTION: How?

The two challenges of convergence are stationarity and mixing point. Mixing and stationarity can be evaluated visually. Common practice is to evaluate stationarity and mixing in trace plots for scalar summaries of interest. In mice these scalar summaries of interest are chain means and chain variances. As Van Buren describes users can also specify a model specific scalar summary (e.g., a regression coefficient). search a user specified scalar summary however is not universal to all complete data problems. Therefore as inspired by someone 2002 we will use introduce a scalar summary that summarizes the multivariate state space of the algorithm. In this study we use the first eigenvalue of the variance covariance matrix of the completed data set per imputation. The methods under evaluation are $\widehat{R}$ and $AC$. These will be computed for the three scalar summaries of interest: chain means, chain variances, and the first eigenvalue of the variance-covariance matrix across imputations.

The two methods will be evaluated with an estimand per scalar summary of interest. It is assumed that when the estimates are unbiased (and confidence valid), the algorithm is sufficiently converged. The $\widehat{R}$ and $AC$ values with chain means as theta are evaluated against the bias in univariate mean estimates (**just for Y??**). Similarly, the performance measure for Rhat and $AC$ applied to chain variances is the bias in estimated standard deviations. To evaluate the performance of Rhat and $AC$ on the eigenvalues, we will use bias in R squared and in the estimated regression coefficient, the coverage rate of the CI95% of the regression estimate, and the CI length. **remove the term performance measure above and add that performance measures are bias in all estimands, and cov and CIL of regression coeff.**

ANSWER

In general there is more bias in the estimate in conditions with a higher proportion of missingness and or a lower number of iterations. As required the values of o$\widehat{R}$ R lower for conditions with a higher number off iterations. End somewhat higher 4 conditions with a higher percentage of missing data. However autocorrelation did not show a decrease with increase number of iterations. After replacing the ACF function with a manual calculation for autocorrelation the autocorrelation values were indeed decreasing with a higher number of iterations. Evaluation with the performance measures shows that are had an autocorrelation are conservative: signs of non-convergence in conditions where estimands are unbiased and or confidence valid. As expected complete convergence (as indicated by autocorrelation equals 0; $\widehat{R}$ equals 1)  is not observed.

QUESTION: When?

@@

ANSWER

Our head is equal to 1.2 is not stringent enough because the performance measures R not unbiased and confidence fellas fell it 4 every condition in which $\widehat{R}$ is smaller than one point 2. At least, not when $\widehat{R}$ is smaller than one point two fur first occurs. The threshold $\widehat{R}$ is equal to one point one is somewhat conservative with respect to the performance measures. It seems OK. $\widehat{R}$ is equal to 1.01 is too stringent compared to the performance measures (**but may be necessary in more complex complete data models???**).

RECOMMENDATIONS

LIMITATIONS/FUTURE RESEARCH

## Start methods section

@@

```{r, include = FALSE}
# load("../Results/test.Rdata")
# source("../2.CreateFigures.R")
```

```{r}
# mean_bias + plot_annotation(title = "Bias in mean estimate, convergence of chain means")
# mean_Rh 
# mean_ACF + plot_annotation(title = "Dotted lines are old $AC$ computation, solid are new.")
```

@@

```{r}
# est_bias + plot_annotation(title = "Bias in regression estimate, coverage rate, \n bias in variance explained")
# est_cov
# Rsq_bias 
```

@@

```{r include=FALSE}
# # chain means
# dat[-1, 6:9] %>% mutate(max_Rh = apply(., 1, max)) %>% .$max_Rh<1.2
# 
# # chain variances
# dat[-1, 26:29] %>% mutate(max_Rh = apply(., 1, max)) %>% .$max_Rh<1.2

@@

>>>>

**Move this after notation??** Figure \ref{fig:diagram} provides an overview of the steps involved with MI---from incomplete data, to $M$ completed datasets, to $M$ estimated quantities of interest ($\hat{Q}$s), to a single pooled estimate $\bar{Q}$. Missing data in $y$ is 'imputed' (i.e., filled in) $M$ times. The imputed data ($y_{imp}$) is combined with the observed data ($y_{obs}$) to create $M$ completed datasets. On each completed dataset, the analysis of scientific interest (the 'substantive model', or 'complete data model') is performed. The quantity of scientific interest (e.g., a regression coefficient) is denoted with $Q$. Since $Q$ is estimated on each completed dataset, $M$ separate $\hat{Q}$-values are obtained. These $M$ values are combined into a single pooled estimate $\bar{Q}$. **ALSO: make clear that iteration happens within the first step.**

```{r diagram, out.width="\\linewidth", include=TRUE, fig.align="center", fig.cap="Scheme of the main steps in multiple imputation."}
knitr::include_graphics("./images/diagram.pdf") #adapted from [@buur18] par. 1.4.1.
```

@@

### Research question: how and when to diagnose non convergence?

Sub-questions: How can non-convergence be diagnosed? Are common MCMC non-convergence diagnostics appropriate for MICE? And if so, which threshold should be used to diagnose non-convergence? How many iterations are sufficient/needed to be able to diagnose non-convergence? Are the default number of iterations sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)? **(These are part of the methods, not RQ per se: What are the effects of continued iterations on estimates, predictions and inferences? And finally, do the answers differ with varying missingness proportions? That is, we vary the nr of iterations and the missingness proportion because we assume that the algorithm has not reached convergence at $t=1$, and performs worse with moderate to high missingness (not so much with little or a LOT of missingness).)**

@@

The two challenges for convergence of iterative algorithms are mixing and stationarity [@gelm13], see Figure \ref{fig:non-conv}. Without mixing, imputation chains do not intermingle nicely. Without stationarity, there is trending within imputation chains. While the current practice is to inspect traceplots for signs of non-mixing and non-stationarity, diagnostic evaluation of these two challenges would be preferred. Non-stationarity within chains may be diagnosed with e.g., autocorrelation [$AC$; @scha97; @gelm13], numeric standard error ['MC error'; @gewe92], or Raftery and Lewis's [-@raft91] procedure to determine the effect of trending on the precision of estimates. A widely used diagnostic to monitor mixing between chains is the potential scale reduction factor $\widehat{R}$ ['Gelman-Rubin statistic'; @gelm92]. With a recently proposed adaptation, $\widehat{R}$ might also serve to diagnose non-stationarity, but this has not yet been thoroughly investigated [@veht19]. Therefore, mixing and stationarity will be evaluated separately in this study. As recommended [e.g., @cowl96, p. 898], $AC$ and $\widehat{R}$ will be used.

```{r non-conv, out.width='.49\\linewidth', fig.width=3.5, fig.height=3,fig.show='hold',fig.cap="Pathological non-convergence versus typical convergence."}
load("./images/nonc.Rdata")
load("./images/conv.Rdata")
par(mar = c(2, 2, 0.1, 0.1))
plot(nonc)
plot(conv)

@@

## How to diagnose non-convergence?

The two challenges of convergence are stationarity and mixing point. Mixing and stationarity can be evaluated visually. Common practice is to evaluate stationarity and mixing in trace plots for scalar summaries of interest. In mice these scalar summaries of interest are chain means and chain variances. As @buur18 describes, users can also specify a model-specific scalar summary (e.g., a regression coefficient). Such a user-specified scalar summary, however, is not universal to all complete data problems. Therefore, as inspired by **someone** (2002), we will use introduce a scalar summary that summarizes the multivariate state space of the algorithm. In this study we use the first eigenvalue of the variance covariance matrix of the completed data set per imputation. The methods under evaluation are $\widehat{R}$ and $AC$. These will be computed for the three scalar summaries of interest: chain means, chain variances, and the first eigenvalue of the variance-covariance matrix across imputations.

The two methods will be evaluated with an estimand per scalar summary of interest. It is assumed that when the estimates are unbiased (and confidence valid), the algorithm is sufficiently converged. The $\widehat{R}$ and $AC$ values with chain means as theta are evaluated against the bias in univariate mean estimates (**just for Y??**). Similarly, the performance measure for Rhat and $AC$ applied to chain variances is the bias in estimated standard deviations. To evaluate the performance of Rhat and $AC$ on the eigenvalues, we will use bias in R squared and in the estimated regression coefficient, the coverage rate of the CI95% of the regression estimate, and the CI length. **remove the term performance measure above and add that performance measures are bias in all estimands, and cov and CI length of regression coeff.**

**Result:** In general there is more bias in the estimate in conditions with a higher proportion of missingness and or a lower number of iterations. As required the values of $\widehat{R}$ are lower for conditions with a higher number off iterations. End somewhat higher in conditions with a higher percentage of missing data. However autocorrelation did not show a decrease with increase number of iterations. After replacing the ACF function with a manual calculation for autocorrelation the autocorrelation values were indeed decreasing with a higher number of iterations. Evaluation with the performance measures shows that are had an autocorrelation are conservative: signs of non-convergence in conditions where estimands are unbiased and or confidence valid. As expected complete convergence (as indicated by autocorrelation equals 0; $\widehat{R}$ equals 1)  is not observed.

## When to diagnose non-convergence?

@@

**Result:** $\widehat{R}$ is equal to 1.2 is not stringent enough because the performance measures R not unbiased and confidence fellas fell it 4 every condition in which $\widehat{R}$ is smaller than one point 2. At least, not when $\widehat{R}$ is smaller than one point two fur first occurs. The threshold $\widehat{R}$ is equal to one point one is somewhat conservative with respect to the performance measures. It seems OK. $\widehat{R}$ is equal to 1.01 is too stringent compared to the performance measures (**but may be necessary in more complex complete data models???**).

## Start old methods section

@@

```{r load-res, include = FALSE}
# load("../Results/test.Rdata")
# source("../2.CreateFigures.R")
```

```{r res-univ}
# mean_bias + plot_annotation(title = "Bias in mean estimate, convergence of chain means")
# mean_Rh 
# mean_ACF + plot_annotation(title = "Dotted lines are old $AC$ computation, solid are new.")
```

@@

```{r res-multi}
# est_bias + plot_annotation(title = "Bias in regression estimate, coverage rate, \n bias in variance explained")
# est_cov
# Rsq_bias 
```

@@

```{r thresh-Rh, include=FALSE}
# # chain means
# dat[-1, 6:9] %>% mutate(max_Rh = apply(., 1, max)) %>% .$max_Rh<1.2
# 
# # chain variances
# dat[-1, 26:29] %>% mutate(max_Rh = apply(., 1, max)) %>% .$max_Rh<1.2

@@

## Recommendations for empirical researchers

## Recommendations for future research

### 2020-04-22 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex
Compute rhat and ac for pathological non-convergence case. Adjust functions by adding wrapper for general convergence function

\textbf{Move this after notation??} Figure 1 provides an overview of the
steps involved with MI---from incomplete data, to \(M\) completed
datasets, to \(M\) estimated quantities of interest (\(\hat{Q}\)s), to a
single pooled estimate \(\bar{Q}\). Missing data in \(y\) is `imputed'
(i.e., filled in) \(M\) times. The imputed data (\(y_{imp}\)) is
combined with the observed data (\(y_{obs}\)) to create \(M\) completed
datasets. On each completed dataset, the analysis of scientific interest
(the `substantive model', or `complete data model') is performed. The
quantity of scientific interest (e.g., a regression coefficient) is
denoted with \(Q\). Since \(Q\) is estimated on each completed dataset,
\(M\) separate \(\hat{Q}\)-values are obtained. These \(M\) values are
combined into a single pooled estimate \(\bar{Q}\). \textbf{ALSO: make
clear that iteration happens within the first step.}

@@

\caption{Scheme of the main steps in multiple imputation.}\label{fig:unnamed-chunk-1}
\end{figure}

@@

Main question: How can non-convergence be diagnosed? Sub-questions: Are
common MCMC non-convergence diagnostics appropriate for MICE? And if so,
which threshold should be used to diagnose non-convergence? How many
iterations are sufficient/needed to be able to diagnose non-convergence?
Are the default number of iterations sufficient (i.e., 5 in mice, 10 in
SPSS and Stata, 30 in mi)? \textbf{(These are part of the methods, not
RQ per se: What are the effects of continued iterations on estimates,
predictions and inferences? And finally, do the answers differ with
varying missingness proportions? That is, we vary the nr of iterations
and the missingness proportion because we assume that the algorithm has
not reached convergence at \(t=1\), and performs worse with moderate to
high missingness (not so much with little or a LOT of missingness).)}

@@

The two challenges for convergence of iterative algorithms are mixing
and stationarity \citep{gelm13}, see Figure XYZ. Without mixing,
imputation chains do not intermingle nicely. Without stationarity, there
is trending within imputation chains. While the current practice is to
inspect traceplots for signs of non-mixing and non-stationarity,
diagnostic evaluation of these two challenges would be preferred.
Non-stationarity within chains may be diagnosed with e.g.,
autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric standard
error \citep[`MC error';][]{gewe92}, or Raftery and Lewis's
\citeyearpar{raft91} procedure to determine the effect of trending on
the precision of estimates. A widely used diagnostic to monitor mixing
between chains is the potential scale reduction factor \(\widehat{R}\)
\citep[`Gelman-Rubin statistic';][]{gelm92}. With a recently proposed
adaptation, \(\widehat{R}\) might also serve to diagnose
non-stationarity, but this has not yet been thoroughly investigated
\citep{veht19}. Therefore, mixing and stationarity will be evaluated
separately in this study. As recommended \citep[e.g.,][p.~898]{cowl96},
\(AC\) and \(\widehat{R}\) will be used.

@@

{\centering \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/unnamed-chunk-2-1} \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/unnamed-chunk-2-2}

@@

\caption{Pathological non-convergence versus typical convergence.}\label{fig:unnamed-chunk-2}
\end{figure}

@@

Research question when and how to diagnose non convergence?

QUESTION: How?

The two challenges of convergence are stationarity and mixing point.
Mixing and stationarity can be evaluated visually. Common practice is to
evaluate stationarity and mixing in trace plots for scalar summaries of
interest. In mice these scalar summaries of interest are chain means and
chain variances. As Van Buren describes users can also specify a model
specific scalar summary (e.g., a regression coefficient). search a user
specified scalar summary however is not universal to all complete data
problems. Therefore as inspired by someone 2002 we will use introduce a
scalar summary that summarizes the multivariate state space of the
algorithm. In this study we use the first eigenvalue of the variance
covariance matrix of the completed data set per imputation. The methods
under evaluation are \(\widehat{R}\) and \(AC\). These will be computed
for the three scalar summaries of interest: chain means, chain
variances, and the first eigenvalue of the variance-covariance matrix
across imputations.

The two methods will be evaluated with an estimand per scalar summary of
interest. It is assumed that when the estimates are unbiased (and
confidence valid), the algorithm is sufficiently converged. The
\(\widehat{R}\) and \(AC\) values with chain means as theta are
evaluated against the bias in univariate mean estimates (\textbf{just
for Y??}). Similarly, the performance measure for Rhat and \(AC\)
applied to chain variances is the bias in estimated standard deviations.
To evaluate the performance of Rhat and \(AC\) on the eigenvalues, we
will use bias in R squared and in the estimated regression coefficient,
the coverage rate of the CI95\% of the regression estimate, and the CI
length. \textbf{remove the term performance measure above and add that
performance measures are bias in all estimands, and cov and CIL of
regression coeff.}

ANSWER

In general there is more bias in the estimate in conditions with a
higher proportion of missingness and or a lower number of iterations. As
required the values of o\(\widehat{R}\) R lower for conditions with a
higher number off iterations. End somewhat higher 4 conditions with a
higher percentage of missing data. However autocorrelation did not show
a decrease with increase number of iterations. After replacing the ACF
function with a manual calculation for autocorrelation the
autocorrelation values were indeed decreasing with a higher number of
iterations. Evaluation with the performance measures shows that are had
an autocorrelation are conservative: signs of non-convergence in
conditions where estimands are unbiased and or confidence valid. As
expected complete convergence (as indicated by autocorrelation equals 0;
\(\widehat{R}\) equals 1) is not observed.

QUESTION: When?

@@

ANSWER

Our head is equal to 1.2 is not stringent enough because the performance
measures R not unbiased and confidence fellas fell it 4 every condition
in which \(\widehat{R}\) is smaller than one point 2. At least, not when
\(\widehat{R}\) is smaller than one point two fur first occurs. The
threshold \(\widehat{R}\) is equal to one point one is somewhat
conservative with respect to the performance measures. It seems OK.
\(\widehat{R}\) is equal to 1.01 is too stringent compared to the
performance measures (\textbf{but may be necessary in more complex
complete data models???}).

RECOMMENDATIONS

LIMITATIONS/FUTURE RESEARCH

@@

>>>>

\textbf{Move this after notation??} Figure \ref{fig:diagram} provides an
overview of the steps involved with MI---from incomplete data, to \(M\)
completed datasets, to \(M\) estimated quantities of interest
(\(\hat{Q}\)s), to a single pooled estimate \(\bar{Q}\). Missing data in
\(y\) is `imputed' (i.e., filled in) \(M\) times. The imputed data
(\(y_{imp}\)) is combined with the observed data (\(y_{obs}\)) to create
\(M\) completed datasets. On each completed dataset, the analysis of
scientific interest (the `substantive model', or `complete data model')
is performed. The quantity of scientific interest (e.g., a regression
coefficient) is denoted with \(Q\). Since \(Q\) is estimated on each
completed dataset, \(M\) separate \(\hat{Q}\)-values are obtained. These
\(M\) values are combined into a single pooled estimate \(\bar{Q}\).
\textbf{ALSO: make clear that iteration happens within the first step.}

@@

\caption{Scheme of the main steps in multiple imputation.}\label{fig:diagram}
\end{figure}

@@

\hypertarget{research-question-how-and-when-to-diagnose-non-convergence}{%
\subsubsection{Research question: how and when to diagnose non
convergence?}\label{research-question-how-and-when-to-diagnose-non-convergence}}

Sub-questions: How can non-convergence be diagnosed? Are common MCMC
non-convergence diagnostics appropriate for MICE? And if so, which
threshold should be used to diagnose non-convergence? How many
iterations are sufficient/needed to be able to diagnose non-convergence?
Are the default number of iterations sufficient (i.e., 5 in mice, 10 in
SPSS and Stata, 30 in mi)? \textbf{(These are part of the methods, not
RQ per se: What are the effects of continued iterations on estimates,
predictions and inferences? And finally, do the answers differ with
varying missingness proportions? That is, we vary the nr of iterations
and the missingness proportion because we assume that the algorithm has
not reached convergence at \(t=1\), and performs worse with moderate to
high missingness (not so much with little or a LOT of missingness).)}

@@

The two challenges for convergence of iterative algorithms are mixing
and stationarity \citep{gelm13}, see Figure \ref{fig:non-conv}. Without
mixing, imputation chains do not intermingle nicely. Without
stationarity, there is trending within imputation chains. While the
current practice is to inspect traceplots for signs of non-mixing and
non-stationarity, diagnostic evaluation of these two challenges would be
preferred. Non-stationarity within chains may be diagnosed with e.g.,
autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric standard
error \citep[`MC error';][]{gewe92}, or Raftery and Lewis's
\citeyearpar{raft91} procedure to determine the effect of trending on
the precision of estimates. A widely used diagnostic to monitor mixing
between chains is the potential scale reduction factor \(\widehat{R}\)
\citep[`Gelman-Rubin statistic';][]{gelm92}. With a recently proposed
adaptation, \(\widehat{R}\) might also serve to diagnose
non-stationarity, but this has not yet been thoroughly investigated
\citep{veht19}. Therefore, mixing and stationarity will be evaluated
separately in this study. As recommended \citep[e.g.,][p.~898]{cowl96},
\(AC\) and \(\widehat{R}\) will be used.

@@

{\centering \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-1} \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-2}

@@

\caption{Pathological non-convergence versus typical convergence.}\label{fig:non-conv}
\end{figure}

@@

\hypertarget{how-to-diagnose-non-convergence}{%
\subsection{How to diagnose
non-convergence?}\label{how-to-diagnose-non-convergence}}

The two challenges of convergence are stationarity and mixing point.
Mixing and stationarity can be evaluated visually. Common practice is to
evaluate stationarity and mixing in trace plots for scalar summaries of
interest. In mice these scalar summaries of interest are chain means and
chain variances. As \citet{buur18} describes, users can also specify a
model-specific scalar summary (e.g., a regression coefficient). Such a
user-specified scalar summary, however, is not universal to all complete
data problems. Therefore, as inspired by \textbf{someone} (2002), we
will use introduce a scalar summary that summarizes the multivariate
state space of the algorithm. In this study we use the first eigenvalue
of the variance covariance matrix of the completed data set per
imputation. The methods under evaluation are \(\widehat{R}\) and \(AC\).
These will be computed for the three scalar summaries of interest: chain
means, chain variances, and the first eigenvalue of the
variance-covariance matrix across imputations.

The two methods will be evaluated with an estimand per scalar summary of
interest. It is assumed that when the estimates are unbiased (and
confidence valid), the algorithm is sufficiently converged. The
\(\widehat{R}\) and \(AC\) values with chain means as theta are
evaluated against the bias in univariate mean estimates (\textbf{just
for Y??}). Similarly, the performance measure for Rhat and \(AC\)
applied to chain variances is the bias in estimated standard deviations.
To evaluate the performance of Rhat and \(AC\) on the eigenvalues, we
will use bias in R squared and in the estimated regression coefficient,
the coverage rate of the CI95\% of the regression estimate, and the CI
length. \textbf{remove the term performance measure above and add that
performance measures are bias in all estimands, and cov and CI length of
regression coeff.}

\textbf{Result:} In general there is more bias in the estimate in
conditions with a higher proportion of missingness and or a lower number
of iterations. As required the values of \(\widehat{R}\) are lower for
conditions with a higher number off iterations. End somewhat higher in
conditions with a higher percentage of missing data. However
autocorrelation did not show a decrease with increase number of
iterations. After replacing the ACF function with a manual calculation
for autocorrelation the autocorrelation values were indeed decreasing
with a higher number of iterations. Evaluation with the performance
measures shows that are had an autocorrelation are conservative: signs
of non-convergence in conditions where estimands are unbiased and or
confidence valid. As expected complete convergence (as indicated by
autocorrelation equals 0; \(\widehat{R}\) equals 1) is not observed.

\hypertarget{when-to-diagnose-non-convergence}{%
\subsection{When to diagnose
non-convergence?}\label{when-to-diagnose-non-convergence}}

@@

\textbf{Result:} \(\widehat{R}\) is equal to 1.2 is not stringent enough
because the performance measures R not unbiased and confidence fellas
fell it 4 every condition in which \(\widehat{R}\) is smaller than one
point 2. At least, not when \(\widehat{R}\) is smaller than one point
two fur first occurs. The threshold \(\widehat{R}\) is equal to one
point one is somewhat conservative with respect to the performance
measures. It seems OK. \(\widehat{R}\) is equal to 1.01 is too stringent
compared to the performance measures (\textbf{but may be necessary in
more complex complete data models???}).

@@

\hypertarget{recommendations-for-empirical-researchers}{%
\subsection{Recommendations for empirical
researchers}\label{recommendations-for-empirical-researchers}}

\hypertarget{recommendations-for-future-research}{%
\subsection{Recommendations for future
research}\label{recommendations-for-future-research}}
